---
title: "Benchmarking Report"
author: "Esha Joshi"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
   rmd: "benchmarking_report.Rmd"
output:
  html_document:
    theme: default
    toc: yes
    toc_depth: 5
---

```{r data.setup, include=FALSE, echo=FALSE, eval=TRUE}
# Load required packages
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)
require(knitr, quietly = TRUE)
require(ggplot2, quietly=TRUE)
require(RColorBrewer, quietly=TRUE)
require(dplyr, quietly = TRUE)
require(gluedown, quietly = TRUE)
require(gridExtra, quietly = TRUE)
require(lubridate, quietly = TRUE)

# set input variable from snakemake
benchmarks.table <- snakemake@input[["benchmarks"]]

# read in tsv
dataset <- read.table(benchmarks.table, header=TRUE, sep="\t")

# some light date/time manipulation for ease in plotting
dataset$times <- hms(dataset$h.m.s)
d.lub <- hour(dataset$times) + minute(dataset$times)/60 + second(dataset$times)/3600
dataset$walltime <- d.lub

# some upfront unit conversions  & derivations for ease in comprehension
# # convert IO units from MB to GB
dataset$io_in <- dataset$io_in / 1000
dataset$io_out <- dataset$io_out / 1000

# # convert CPU time from seconds to hours
dataset$cpu_time_hrs <- dataset$cpu_time / 3600

# # calculate cpu/walltime ratio
dataset$cpuwall_ratio <- (dataset$cpu_time / dataset$s)

# subset the dataset to get data for rules where there is more than 1
# observation/process to pass to geom_boxplot
subset.df <- dataset %>% group_by(rule) %>% filter(n() > 1)

# subset the dataset to get data points for rules meeting time_threshold
# parameter specified at the rule level
time.param <- snakemake@params[["threshold"]]
time.threshold <- time.param * 60
filtered.df <- dataset %>% group_by(rule) %>% filter(s >= time.threshold)

# cluster config version
cluster.config <- snakemake@params[["clusterVersion"]]
```

```{r plot.functions, eval=TRUE, echo=FALSE}
# plotting function for box/jitter plot where single data points are omitted
# from boxplots
create.boxjitter.plot <- function(df, df.subset, xvar, yvar, xlab, ylab) {
  bd.plot <- ggplot(data=df, aes_string(x=xvar, y=yvar)) +
    geom_boxplot(data = df.subset, aes_string(fill=xvar), show.legend = FALSE)  +
    geom_point(position="identity", show.legend = FALSE, alpha=1/2, size=1) +
    labs(x=xlab, y=ylab) +
    theme_minimal() +
    theme(axis.text.x=element_text(angle = 90, vjust=0.5, hjust=1),
          axis.title = element_text(size = 14),
          axis.text = element_text(size = 12))
  bd.plot
}

create.subset.plot <- function(subdf, xvar, yvar, xlab, ylab, title){
  sb.plot <- ggplot(data = subdf, aes_string(x = xvar, y = yvar)) +
    geom_boxplot(aes_string(fill=xvar), show.legend = FALSE) +
    geom_point(position = "identity", show.legend = FALSE, alpha=1/2, size=1) +
    labs(x=xlab, y=ylab, title=title) +
    theme_minimal() +
    theme(axis.text.x=element_text(angle = 90, vjust=0.5, hjust=1),
          axis.title = element_text(size = 14),
          axis.text = element_text(size = 12))
  sb.plot
}
```

Cluster version used: `r cat(cluster.config)`

This is a (development version) R Markdown report that can generated for the pipeline with visualizations for the specified rules and their performance benchmarks.

## Resource Usage and Run Statistics Overview

#### Rules included in this report: (`r length(unique(dataset[["rule"]]))`)
```{r list.rules, echo=FALSE, include=TRUE, results='asis'}
inlines <- c(
  md_bold(unique(dataset$rule))
)
md_bullet(inlines)
```
&nbsp;
&nbsp;
&nbsp;

#### Overview of All Rules

##### Execution time
Total elapsed time displayed in hours for all rules in this workflow.

```{r all.time.exec, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
# plot walltime for all rules benchmarked
exec.time.all <- create.boxjitter.plot(dataset, subset.df, "rule", "walltime", "Rule", "Total walltime in hours for all rules")
exec.time.all
```

#### Memory {.tabset .tabset-pills}
These plots display memory usage for **all rules** in the form of maximum "Resident Set Size", "Virtual Memory Size",
"Unique Set Size", and "Proportional Set Size" in MB. These metrics are derived from Snakemake's [benchmark](https://snakemake.readthedocs.io/en/stable/tutorial/additional_features.html) directive which uses
functionalities provided by the [psutil](https://psutil.readthedocs.io/en/latest/) (python system and process
utilities) library.

##### PSS
Proportional Set Size: The amount of memory occupied for a particular process plus the shared memory across one or more other processes.
```{r max.pss.all, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
max.pss.all <- create.boxjitter.plot(dataset, subset.df, "rule", "max_pss",
                                     "Rule", "PSS (MB) for all rules") +
  scale_y_continuous(n.breaks = 12)
max.pss.all
```

##### USS
Unique Set Size: The amount of memory occupied by a process, unshared and unique to only that process.
```{r max.uss.all, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
max.uss.all <- create.boxjitter.plot(dataset, subset.df, "rule", "max_uss",
                                     "Rule", "USS (MB) for all rules") +
  scale_y_continuous(n.breaks = 12)
max.uss.all
```

##### RSS
Resident Set Size: The amount of memory non-swapped physical memory occupied by a process.
```{r max.rss.all, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
max.rss.all <- create.boxjitter.plot(dataset, subset.df, "rule", "max_rss",
                                     "Rule", "RSS (MB) for all rules") +
  scale_y_continuous(n.breaks = 12)
max.rss.all
```

##### VMS
Virtual Memory Size: The amount of virtual memory occupied by a process.
```{r max.vms.all, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
max.vms.all <- create.boxjitter.plot(dataset, subset.df, "rule", "max_vms",
                                     "Rule", "VMS (MB) for all rules") +
  scale_y_continuous(n.breaks = 12)
max.vms.all
```


### Resource Usage Detailed View
Displays snakemake's benchmark metrics for rules meeting the user-specified time threshold provided at the config level.

```{r threshold.stat, echo=FALSE, results='asis'}
cat(paste0("Time threshold (minutes): ", time.param))
```

&nbsp;

#### Memory {.tabset .tabset-pills}
These plots display memory usage in the form of maximum "Proportional Set Size (PSS)", "Unique Set Size (USS)", "Resident Set Size (RSS)", "Virtual Memory Size (VMS)" for rules that **only** meet the user-specified time threshold.

##### PSS
Proportional Set Size: The amount of memory occupied for a particular process plus the shared memory across one or more other processes.
```{r max.pss.detailed, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
# plot the memory usage with boxplots for rules with >1 process otherwise
# just plot the datapoint
max.pss.detail <- create.boxjitter.plot(filtered.df, subset.df, "rule", "max_pss", "Rule", "PSS (MB)") +
  scale_y_continuous(n.breaks = 12)
max.pss.detail
```

##### USS
Unique Set Size: The amount of memory occupied by a process, unshared and unique to only that process.
```{r max.uss.detailed, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
max.uss.detail <- create.boxjitter.plot(filtered.df, subset.df, "rule", "max_uss", "Rule", "USS (MB)") +
  scale_y_continuous(n.breaks = 12)
max.uss.detail
```

##### RSS
Resident Set Size: The amount of memory non-swapped physical memory occupied by a process.
```{r max.rss.detailed, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
max.rss.detail <- create.boxjitter.plot(dataset, subset.df, "rule", "max_rss", "Rule", "RSS (MB)") +
  scale_y_continuous(n.breaks = 12)
max.rss.detail
```

##### VMS
Virtual Memory Size: The amount of virtual memory occupied by a process.
```{r max.vms.detailed, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
max.vms.detail <- create.boxjitter.plot(filtered.df, subset.df, "rule", "max_vms", "Rule", "VMS (MB)") +
  scale_y_continuous(n.breaks = 12)
max.vms.detail
```
##### Detailed view
Rules using relatively high or low amounts of memory (based on max pss).
```{r mem.det.grid, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
# plot detailed grid view
mem.df1 <- subset(filtered.df, max_pss > 5000)
mem.df2 <- subset(filtered.df, max_pss < 5000)

max.pss.subset1 <- create.subset.plot(mem.df1, "rule", "max_pss", "Rule", "max_pss (MB)", "Rules using memory >5GB")
max.pss.subset2 <- create.subset.plot(mem.df2, "rule", "max_pss", "Rule", "max_pss (MB)", "Rules using memory <5GB")
grid.arrange(max.pss.subset1, max.pss.subset2, nrow=1)
```



#### I/O {.tabset .tabset-pills}
The amount of information processed by the system through I/O (input-output) operations.

##### Read
The amount of information cumulatively read by the system for each process in MB.
```{r, i.read, echo=FALSE, eval=TRUE, fig.align='center', message=FALSE, warning=FALSE}
# plot the io read
io.read <- create.boxjitter.plot(filtered.df, subset.df, "rule", "io_in", "Rule", "Cumulative GB Read")
io.read
```

##### Write
The amount of information cumulatively written by the system for each process in MB.
```{r, i.write, echo=FALSE, eval=TRUE, fig.align='center', message=FALSE, warning=FALSE}
# plot io write
io.write <- create.boxjitter.plot(filtered.df, subset.df, "rule", "io_out", "Rule", "Cumulative GB Written")
io.write
```

##### Detailed view
```{r io.grid, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
# plot detailed grid view
ioin.df1 <- subset(filtered.df, io_in > 10)
ioin.df2 <- subset(filtered.df, io_in < 10)

ioout.df1 <- subset(filtered.df, io_out > 10)
ioout.df2 <- subset(filtered.df, io_out < 10)

ioin.subset1 <- create.subset.plot(ioin.df1, "rule", "io_in", "Rule", "Cumulative GB Written", "Rules with >10GB Read")
ioout.subset1 <- create.subset.plot(ioout.df1, "rule", "io_out", "Rule", "Cumulative GB Written", "Rules with >10GB Written")

grid.arrange(ioin.subset1, ioout.subset1, nrow=2)
```

#### CPU Time {.tabset .tabset-pills}

##### Total CPU time
The total CPU time is the amount of time for which a CPU was used to process information performing logic or arithmetic operations for a particular process (displayed in hours).

```{r, cpu.time, eval=TRUE, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
# plot the CPU time in hours
cpu.time <- create.boxjitter.plot(filtered.df, subset.df, "rule", "cpu_time_hrs", "Rule", "CPU time in hours")
cpu.time
```

##### CPU to Walltime Ratio
Programs and applications do not use the CPU or processor 100% of the time and some time is spent on I/O operation, waiting for the network or disk. The relationship of CPU time to walltime (or CPU/second) for a process can provide insight on potential bottlenecks with CPU utilization, network bandwidth, disk contention, system overhead, etc. More information on this can be found [here](https://pythonspeed.com/articles/blocking-cpu-or-io/).

For **single-threaded** processes:\
CPU/s ≈ 1 indicates the process spent all of its time using the CPU.\
CPU/s < 1 indicates the process spent 1-(CPU/s) of the time waiting.

For **multi-threaded processes** (CPU/s can be as high as *N* CPUs)\
CPU/s < *N* indicates the process spent much of its time waiting.\
CPU/s ≈ *N* indicates the process saturated the allocated number of processors (i.e. for a process with 2 threads, a CPU/s = 2 indicates saturation)

```{r, cputowalltime, eval=TRUE, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
# plot cpu to walltime ratio (i.e. CPU/s)
cpuwall.ratio <- create.boxjitter.plot(filtered.df, subset.df, "rule", "cpuwall_ratio", "Rule", "CPU/second")
cpuwall.ratio
```
