---
title: "WGS Run Summary"
author:
  - "Bari Ballew"
  - "Cameron Palmer"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
  rmd: "run_summary.Rmd"
output:
  html_document:
    code_folding: "hide"
  highlight: tango
  number_sections: no
  theme: default
  toc: yes
  toc_depth: 3
  toc_float:
    collapsed: no
    smooth_scroll: yes
---

```{r load.packages, eval=TRUE, echo=FALSE, message=FALSE}
## Load required R packages

library(knitr, quietly = TRUE)
library(kableExtra, quietly = TRUE)
library(ggplot2, quietly = TRUE)
library(RColorBrewer, quietly = TRUE)
```

```{r load.resources, eval=TRUE, echo=FALSE, message=FALSE}
## Load tested functions for this report
source(snakemake@input[["r_resources"]])
```

```{r set.ggplot2.theme, eval=TRUE, echo=TRUE}
## Configure standard theme information for ggplot2

my.theme <- theme_light() + theme(plot.title = element_text(size = 15, hjust = 0.5),
                                  axis.title = element_text(size = 14),
                                  axis.text = element_text(size = 12),
                                  strip.background = element_blank(),
                                  strip.text = element_text(size = 14, colour = "black"))
```

```{r load.snakemake.files, eval=TRUE, echo=FALSE}
input.subjects <- snakemake@params[["input_samples"]]
fastqc.filename <- snakemake@input[["fastqc"]]
tsv.path <- snakemake@params[["out_prefix"]]
start.time.filename <- snakemake@input[["start_time"]]
```

## Initial QC Subject Tracking

A total of `r length(input.subjects)` subjects were queued up for analysis in this run.  The table below reports lane/read failures for select FastQC metrics.  Only failing fastqs are reported in those columns.  Please see the MultiQC report for additional information.

```{r report.before.after.subjects, eval=TRUE, echo=FALSE, results="asis"}
res <- prepare.initial.subject.tracking.table(input.subjects)
res <- add.fastqc.data(res, fastqc.filename)
res <- res[order(res[, "Subject"]), ]
rownames(res) <- NULL

knitr::kable(res, caption = "Summary of Subject Fate") %>% kableExtra::kable_styling("striped", position = "left", full_width = FALSE)

write.output.table(res, tsv.path)
```

***
<br>

## Total Elapsed Analysis Run Time

```{r report.time, eval=TRUE, echo=FALSE, results="hide"}
start.time <- read.table(start.time.filename, header = FALSE)[1,1]
end.time <- as.integer(format(Sys.time(), "%s"))
elapsed.time <- round((end.time - start.time) / 3600, 2)
```

Elapsed time was `r elapsed.time` hours.

***
<br>

## Assorted Software Links

<br>

***
<br>

## Session Information

<br>

The following summarizes the loaded R configuration for the run that created this report.

```{r session.info, eval=TRUE, echo=TRUE}
sessionInfo()
```
